---
title: "coefficient plots, etc."
output:
  ioslides_presentation
bibliography: ../vis.bib
---

<!-- 
apa.csl is a slightly hacked version of APA 
  (modified for "et al" after 2 authors in text)
-->
<!-- .refs is style for reference page (small text) -->
<style>
.refs {
   font-size: 16px;
}
h2 { 
 color: #3399ff;		
}
h3 { 
 color: #3399ff;		
}
.title-slide {
   background-color: #55bbff;
}
</style>
<!--    content: url(https://i.creativecommons.org/l/by-sa/4.0/88x31.png)
>
<!-- Limit image width and height -->
<style type="text/css">
img {     
  max-height: 530px;     
  max-width: 800px; 
}
</style>

# setup

## load packages

```{r pkgs,message=FALSE}
library(tidyverse)
## graphics
theme_set(theme_bw()+theme(panel.spacing=grid::unit(0,"lines")))
## modeling/coef plots
library(lme4)
library(broom)
library(broom.mixed)
library(dotwhisker)
library(stargazer) ## coef tables
## predictions/effects plots
library(emmeans)
library(effects)
library(ggeffects)
## also: margins, marginaleffects, sjPlot, ...
library(remef) ## remotes::install_github("hohenstein/remef")
library(faux)
```

# coefficient plots

## principles

- make it easy to compare magnitudes
- use appropriate scales
- which comparisons do you want to make?
- typically drop intercept (why?)

## nuts and bolts

- old and busted: `arm::coefplot`, `coefplot2`
- new hotness:
    - `broom[.mixed]` + `ggplot2`
    - `dotwhisker`


## example

```{r contr_fit,cache=TRUE,warning=FALSE}
data(Contraception,package="mlmRev")
Contraception <- Contraception %>%
    mutate(ch = factor(livch != 0, labels = c("N", "Y")))
m3 <- glmer(use ~ age * ch + I(age^2) + urban + (1 | urban:district),
            data=Contraception, family=binomial)
```

## coefficient table {.smaller}

```{r stargazer_out,results="asis"}
stargazer(m3,type="html")
```

**Also see** `broom`/`broom.mixed` + `huxtable` packages

## coefficient plot

```{r dw1}
gg0 <- dotwhisker::dwplot(m3, by_2sd=TRUE)
print(gg0)
```

## add reference line

```{r dw_refline}
gg1 <- gg0 + geom_vline(xintercept=0,lty=2)
print(gg1)
```

## scaling

- need to scale parameters appropriately to compare magnitude
- @gelman_scaling_2008; @schielzeth_simple_2010
- `arm::standardize`, `arm::rescale` (handy but inefficient)

> binary.inputs: options for standardizing binary variables, default is ‘center’; ‘0/1’ keeps original scale; ‘-0.5,0.5’ rescales 0 as -0.5 and 1 as 0.5; ‘center’ subtracts the mean; and ‘full’ subtracts the mean and divides by 2 sd.

- mixed models: standard deviations have same scale as corresponding parameter

## dwplot without auto-scaling

Look at results if we turn *off* auto-scaling (default)
```{r unscaled}
dotwhisker::dwplot(m3, by_2sd = FALSE)
```

## alternative: explicitly scale age parameter

```{r update, cache=TRUE}
Contraception <- Contraception %>%
    mutate(sc_age = drop(scale(age)))
m3_sc <- update(m3, 
      . ~ sc_age * ch + I(sc_age^2) + urban + (1 | urban:district))
```

## plot

```{r dwplot2, warning=FALSE}
dotwhisker::dwplot(m3_sc, effects="fixed", by_2sd = FALSE) +
    geom_vline(xintercept=0,lty=2)
```


## other alternatives

- scaling makes units harder to interpret
- separate sets of predictors into facets according to their units
- ??

## manually tidying

```{r dwplot3}
cc <- broom.mixed::tidy(m3_sc,effects="fixed")
print(cc,digits=3)
```

## GLM for comparison {.smaller}


```{r m3_glm}
m3_fixed <-  glm(
    use ~ sc_age * ch + I(sc_age^2) + urban,
    data=Contraception, family=binomial)
dotwhisker::dwplot(list(sc_GLMM=m3_sc, GLM=m3_fixed))+
    geom_vline(xintercept=0,lty=2)+ scale_colour_brewer(palette="Dark2")
```

## dotwhisker limitations

- sometimes gets "confused" with complex models
- might want to post-process output of `tidy()` (edit variable/term names, etc.)
- use `purrr::map_dfr(list(mod1=, mod2=, .id = "model")` to process a list of models and collapse to a tibble

## alternative: data prep

```{r alternative1}
m3_res <- (purrr::map_dfr(list(with_re = m3_sc, no_re=m3_fixed),
                  tidy,
                  conf.int = TRUE,
                  .id = "model")
  %>% mutate(term=fct_inorder(term))
  %>% filter(term!="(Intercept)")
)
```

(`purrr::map_dfr` == "run function on elements of a list, bind results into a single data frame, add `.id` column based on names of the list elements")

## alternative: plot {.smaller}

```{r altplot,fig.height=3.5, warning=FALSE}
pd <- position_dodge(width=0.5)
(gg5 <- ggplot(m3_res,aes(x=estimate,y=term,colour=model))
  + geom_pointrange(aes(xmin=conf.low,xmax=conf.high),
                    position=pd)
  + labs(y="")
  + scale_colour_brewer(palette="Dark2") + geom_vline(xintercept=0,lty=2))
```

## reorder terms

```{r alt_reorder,fig.height=3.5, warning=FALSE}
m3_res_order <- mutate(m3_res,term=reorder(term,estimate))
gg5 %+% m3_res_order
```

## caterpillar plots

- Useful when plotting *many* parallel/similar coefficients
- e.g.:
   - random effects in a mixed model
   - bioinformatics: genes, microbial species, etc..

## caterpillar plots (utility)

Reorders levels, allowing for multiple terms per level

```{r cat1}
caterpillar_levels <- function(x, order_term = "(Intercept)") {
    lev_order <- (x
        %>% filter(term==order_term)
        %>% group_by(level)
        %>% summarise(across(estimate, mean, na.rm = TRUE))
        %>% arrange(estimate)
        %>% pull(level)
    )
    x <- dplyr::mutate(x, across(level, factor, levels = lev_order))
    return(x)
}
```

## caterpillar plot {.smaller}

```{r cat2, fig.width=12, fig.height=4} 
cat_data  <- (m3
  %>% broom.mixed::tidy(m3, effects="ran_vals")
  %>% select(level, term, estimate, std.error)
  %>% caterpillar_levels()
)
ggplot(cat_data, aes(level, estimate,
                     ymin = estimate-2*std.error,
                     ymax = estimate+2*std.error)) +
  geom_pointrange() +
  guides(x = guide_axis(n.dodge = 3))
```

## summary

- useful in comparing lots of models;  
models with many parallel parameters
- separate parameter types with facets
- sort by magnitude of estimate

## other thoughts

- can often use a wide aspect ratio (don't need much height)
- may want to replace/improve coefficient names (`faux` package?)
- Bayesian models: violin plots of posterior samples? Or 50% + 95% intervals?

# marginal/effects/prediction plots

## emmeans

```{r}
e1 <- emmeans(m3, ~age*ch, at=list(age=-14:20), type="response")
plot(e1)
```

## as.data.frame → ggplot {.smaller}

```{r}
e1d <- as.data.frame(e1)
print(gg1 <- ggplot(e1d,
              aes(age, prob, colour=ch, fill=ch,
                  ymin=asymp.LCL, ymax=asymp.UCL)) +
  geom_line() +
  geom_ribbon(colour=NA, alpha=0.2))
```
  
## effects

```{r effects1, fig.width=10, fig.height=6}
plot(ae <- allEffects(m3, xlevels = list(age=51)))
```

## as.data.frame

```{r effects_asdf}
aed <- as.data.frame(ae)[["age:ch"]]
gg1 %+% aed + aes(y=fit, ymin=lower, ymax=upper)
```

## ggeffects


```{r ggeffects}
ag <- ggeffect(m3, terms=c("age [all]", "ch"))
plot(ag)
```

## partial effects removal

'Correct' data for non-focal terms:

$$
\beta_0 + \sum_{i \in \substack{\textrm{focal} \\ \textrm{params}}} \beta_i x_i + \textrm{resid}_i
$$

or

$$
\textrm{predict}_i - \sum_{i \in \substack{\textrm{non-focal} \\ \textrm{params}}}  \beta_i x_i
$$


## example

```{r remef1, message=FALSE}
fm1 <- lmer(Reaction ~ 1 + Days + (1 + Days | Subject), sleepstudy)
ag <- ggeffect(fm1, terms="Days [all]")
gg2 <- plot(ag) + geom_point(data=sleepstudy,
                             aes(x=Days, y = Reaction, colour=Subject)) +
  colorspace::scale_colour_discrete_qualitative(guide="none")
```

## example

```{r remef2, echo=FALSE}
print(gg2)
```

## example (cont.) {.smaller}

```{r remef3, message=FALSE}
sleepstudy$re <- remef(fm1, ran = list(Subject = "Days"))
gg3 <- plot(ag) + geom_point(data=sleepstudy,
                      aes(x=Days, y = re, colour=Subject)) +
  colorspace::scale_colour_discrete_qualitative(guide="none")
cowplot::plot_grid(gg2, gg3)
```

## faux: nice contrast names

```{r faux}
mm <- (mtcars
  %>% mutate(across(cyl, contr_code_treatment))
)
m1 <- lm(mpg ~ disp*cyl, mm)
dwplot(m1, by_2sd=TRUE) + geom_vline(xintercept = 0, lty=2)
```

## references {.smaller}
