---
title: "high-dimensional data (focus on dynamical views)"
bibliography: ../vis.bib
date: 29 Oct 2021
---

## packages

Some of these packages may be hard to install, i.e. depending on system resources (especially `rggobi`).

```{r pkgs, message=FALSE}
library(tourr)
library(rggobi)
library(factoextra)
library(tidyverse)
library(ggalt)
library(rgl)
```

(`ggobi` has been *archived*: use `remotes::install_version("rggobi", "2.1.22")`
or `remotes::install_github("ggobi/rggobi")`; see the github repository https://github.com/ggobi/rggobi
and the ggobi web site http://ggobi.org/)


## @buja_interactive_1996

- scatterplots (point locations)
- traces (functions: parallel coordinates, [Andrews plots](https://en.wikipedia.org/wiki/Andrews_plot))
- glyphs (stars/radar charts/faces/etc.)

## tasks

* finding gestalt: focusing views (projection)
* posing queries: brushing/linking
* making comparisons: arranging many views

## focusing: animation/motion

* adds a dimension (2D to 3D)
* smoothness/object continuity
* real-time user control: allows user to search/tune

## interactive example

```{r ggobi_laser,eval=FALSE}
ggobi(laser)
```

## tours

three components:

- data matrix ($n \times p$)
- **tour path** that produces a smooth sequence of projection matrices ($p \times d$) (`apropos("_tour$")`)
- **display method** that renders the projected data (`apropos("animate_")`)

optional extras (for "guided tour"):

- search strategy (`apropos("^search_")`)
- index (objective) function (`cmass`, `holes`, `lda_pp`)

Goal: allow automated, systematic search through **entire** space of projections

---

```{r tour_flea, eval=FALSE}
f <- flea[, 1:6]
animate(f,
        tour_path = guided_tour(holes()),
        display = display_xy(col = flea$species))
```

## available display methods

```{r animate_methods}
apropos("animate_")
apropos("_tour$")
```

## PCA adjustment

Examples/data from @holmes_modern_2019

- `prcomp` (SVD): preferred (also `princomp`)
- same data; projects in a way that (hopefully) focuses on interesting angles in the data
- assumes linear features
- normality not necessary, but better-behaved with normal data

## wine PCA example

```{r wine1}
load("../data/wine.RData")      ## numeric matrix (wine)
load("../data/wineClass.RData") ## metadata (wine.class)

winePCAd <- prcomp(wine, scale = TRUE) ## compute PCA
f0 <- fviz_pca_biplot(winePCAd, geom = "point", habillage = wine.class,
                      col.var = "violet", addEllipses = TRUE, ellipse.level = 0.69)
tmp_plot <- function(p) { ## tweaks
  p + ggtitle("") + coord_fixed() +
    scale_colour_brewer(palette="Dark2") +
    scale_fill_brewer(palette="Dark2")
}
```

## plot

```{r plot_wine1}
print(tmp_plot(f0))
```

## get augmented data

```{r aug_data}
aug_data <- (get_pca_ind(winePCAd)$coord
  %>% as_tibble()
  %>% mutate(class = wine.class, .before=1)
)
```

## improved plot

```{r plot_wine2,message=FALSE}
f1 <- fviz_pca_biplot(winePCAd, geom = "point", habillage = wine.class,
   col.var = "violet")
f2 <- tmp_plot(f1) + ggalt::geom_encircle(data = aug_data,
                                aes(x=Dim.1,y=Dim.2,group=class,
                                    colour=class,
                                    fill=class),alpha=0.2,
                                show.legend=FALSE)
```

--- 

```{r print_wineplot2}
print(f2)
```

```{r}
adjust_pcaplot <- function(p, pca, dims=1:2) {
  sdvec <- pca$sdev[dims]
  p + coord_fixed(ratio = sdvec[2]/sdvec[1])
}
f3 <- adjust_pcaplot(f2, winePCAd)
```

---

```{r print_wineplot3}
print(f3)
```

```{r rgl, eval=FALSE}
cvec <- RColorBrewer::brewer.pal("Dark2", n=3)
with(aug_data,
     plot3d(Dim.1, Dim.2, Dim.3, col=cvec[wine.class], size=4))
with(aug_data,
     plot3d(Dim.1, Dim.2, Dim.3, col=cvec[wine.class], size=2, type = "s"))
```

## references

